---
title: "Abandoned wells data exploration"
author: Ronny A. Hernandez Mora
execute:
  message: false
  warning: false
format: 
  html:
    theme:
      - flatly
    linkcolor: "#FF5500"
    highlight-style: tango
    toc: true
    toc-title: Table of contents
    toc-location: left
    number-sections: false
    colorlinks: true
    code-fold: true
    code-line-numbers: true
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
library(dplyr)
library(sf)
library(janitor)
library(purrr)
library(visdat)
library(ggplot2)
library(rmarkdown)
```

This is an Exploratory Data Analysis to understand the data characteristics of 
the layer `o16_WellsAbnd_HFI_2021` from the [Wall-to-Wall Human Footprint Inventory](https://abmi.ca/home/data-analytics/da-top/da-product-overview/Human-Footprint-Products/HF-inventory.html) by the [Alberta Biodiversity Monitoring Institute](https://abmi.ca/home.html).

The aim is to identify characteristics that may aid in 
[selecting polygons](https://github.com/ronnyhdez/reclaimed_sites_ab/issues/6)
for subsequent analysis.

# Read the data

Following information is part of the file metadata:

```{r}
abandoned_wells <- st_read(dsn = 'HFI2021.gdb/HFI2021.gdb',
                           layer = 'o16_WellsAbnd_HFI_2021') |> 
        clean_names()
```

This is a dataframe with `r nrow(abandoned_wells)` observations and 
`r ncol(abandoned_wells)` variables. A glimpse of the first observations per
each variable and its type is shown here:

```{r}
glimpse(abandoned_wells)
```

# Attribute table general characteristics

Use just the attribute table as dataframe for quick exploration

```{r}
abandoned_wells_df <- abandoned_wells |> 
        st_drop_geometry()
```

How many unique values do we have per variable?

```{r}
abandoned_wells_df |> 
        map_dfr(~n_distinct(.x)) |> 
        paged_table()
```

Insights from the output:

 - `footprint_id` does not match the number of observations. Check if there are
 NA's in polygons and why.
 - `well_status` we just have 1 which means **abandoned**.
 - `spud_date` and `spud_year` are the date when drilling first began on the
 well pad. 125 years in the registry? Check how many years of data we have.
 - The `number_wellheads` is the total number of wellbores on the well pad. Does
 this have an impact in the reclamation process?
 - `visible` variable probably have `NA's`. There should be just 3 values:
 0 = not visible, 1 = fully visible, 2 = partially visible. This reflects the
 state of visibility of a footprint. No idea what this means.
 - `oilsands_evaluation_well` we have just the 2 categories described in
 documentation: 0 = No Oil Sands Evaluation (OSE) well, 1 = OSE well present.
 - Check distribution of variable `reclamation_status`.
 - `reclamation_date` it's the most recent reclamation certification.

## Character variables

```{r}
abandoned_wells_df |> 
        select_if(is.character) |> 
        glimpse() 
```

 - We have just 6 character variables. 
 - Take into account that other variables with categories are encoded with
 numbers and I should check the ABMI data document to understand the meaning.

Check unique values per character variable:

```{r}
# Check number of distinct observations
abandoned_wells_df |> 
        select_if(is.character) |> 
        map(~n_distinct(.x))

# Check categories 
abandoned_wells_df |> 
        select_if(is.character) |> 
        select(-hfi_id, -disp_num, -lcu_id) |>
        map(~unique(.x))
```

```{r}
# abandoned_wells_df |> 
#         vis_dat(warn_large_data = FALSE)
```

# Date variables

We have the following variables that represents a date in the dataset:

| Variable | Meaning |
| -------- | ------- |
| first_spud_date            | date when drilling first began on the well pad
| first_spud_year            | date when drilling first began on the well pad
| reclamation_date           | most recent reclamation certification issued
| min_onproduction_date      | earliest reported year a well went into prod
| max_last_production_date   | latest reported year of production at a well
| max_abandoned_date         | latest reported year a well was abandoned
| year                       | ? (Did not find definition in documentation)
 
We need to obtain the latest reclamation date for each of the polygons. One
site could potentially have several reclamation dates, but we have no way to
find out that information from this dataset. Also we are considering using just
the last reclamation date for the analysis. This dataset has reclamation
dates that span `r range(abandoned_wells_df$reclamation_date)`.

```{r}
abandoned_wells_df |> 
        group_by(reclamation_date) |> 
        tally() |> 
        paged_table()

zero_obs <- abandoned_wells_df |> 
        group_by(reclamation_date) |> 
        tally() |> 
        filter(reclamation_date == 0) |> 
        select(n) |> 
        pull()
```

::: {.callout-important}
From the summarized table, there are `r zero_obs` observations with a year
reclamation **date value of 0**. This means that if reclamation date is
an important variable for the analysis, we will loose these polygons.
:::

Let's check the number of sites per year of the most recent reclamation
certification issued:

```{r}
abandoned_wells_df |> 
        group_by(reclamation_date) |> 
        tally() |> 
        filter(reclamation_date != 0) |> 
        ggplot(aes(x = reclamation_date, y = n)) +
        geom_col() +
        labs(x = "Year of the most recent reclamation certification issued",
             y = "Total well sites") +
        theme_light()
```

## Number of observations with minimun amount of years from reclamation year

We need to obtain polygons with a sufficient time range from the last
reclamation year to include in the analysis. The following interactive plot
helps to explore the number of observations that we have if we select the
number of years since the last reclamation certification was issued.

::: {.callout-important}
The total number of observations from the dataset is `r nrow(abandoned_wells_df)`
but there are `r zero_obs` observations with a year reclamation **date value of 0**. 
:::

```{r}
# Create interactive plot that shows number of observations with at least
# 20 years or X since last year of reclamation



```

## Reclamation date vs Abandoned date

Does the reclamation date has to match with abandoned date?

Differences would mean: 

0 = Same year abandoned date reclamation.
> 0 = Date reclamation is more recent than the abandoned date.
< 0 = Abandoned date is more recent than reclamation date.

```{r}
abandoned_wells_df |> 
        filter(reclamation_date != 0 & max_abandoned_date != 0) |> 
        mutate(diff = reclamation_date - max_abandoned_date) |> 
        select(diff, reclamation_date, max_abandoned_date) |> 
        # arrange(desc(diff))
        # paged_table()
        group_by(diff) |> 
        tally() |> 
        # paged_table()
        ggplot(aes(x = diff, y = n)) +
        geom_col() +
        theme_light()
```

## Reclamation date vs Last Production date

Here I have the same question: Does the reclamation date has to match with
the last production date? Given that we need polygons with no buildings or 
infrastructure on it to avoid anomalies in the spectral signature 

## Observations

# Key points:

 - We do have the last reclamation year
 - There are xxx NA's